---
title: "class11"
author: "Vaibhav Menon"
date: "5/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##PDB Statistics

Download CSV file from PDB database <http://www.rcsb.org/stats/summary>. Read this into R and determine fraction of X-ray structures.
```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
View(pdb.stats)
```

Let's calculate something
```{r}
percent <- (pdb.stats$Total /sum(pdb.stats$Total))*100
names(percent) <- pdb.stats$Experimental.Method
percent

proteintot <- (sum(pdb.stats$Proteins) /sum(pdb.stats$Total))*100
proteintot

```
Q1: X-ray crystallography accounts for 89.5% of PDB entries. Proteins account for 92.8% of total structures.

Q2: There are 113 protease structures in the results, which account for 2.6% of the structures.

# Working with Bio3d

Let's load the package
```{r}
library(bio3d)
```

List the functions in the package
```{r}
lbio3d()
```

Read a pdb file
```{r}
pdb <- read.pdb("1hsg")
pdb
```

Select and write-out C-alpha atoms only.
```{r}
inds <- atom.select(pdb, "calpha")
inds
```

check if we have what we want
```{r}
head(pdb$atom[inds$atom,])
```

```{r}
ca.pdb <- trim.pdb(pdb, inds=inds)
write.pdb(ca.pdb, file="calpha_only.pdb")
```

# Download some example PDB files 
```{r}
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
# Extract and align the chains we are interested in  
pdbs <- pdbaln(files, fit = TRUE)
# Print to screen a summary of the 'pdbs' object
pdbs
```
```{r}
# Access the first 5 rows, and 8 columns
pdbs$ali[1:5, 1:8]

```

```{r}
# Associated residues numbers
pdbs$resno[1:5, 1:8]
```

```{r}
# Inspect alignment for gaps
gaps <- gap.inspect(pdbs$ali)   

# Indices of non-gap columns
head(gaps$f.inds) 
 
# Access non-gap columns
pdbs$ali[, gaps$f.inds]
 
```
```{r}
# Inspect alignment for gaps
gaps <- gap.inspect(pdbs$ali)   

# Indices of non-gap columns
head(gaps$f.inds) 
 
# Access non-gap columns
pdbs$ali[, gaps$f.inds]
basename.pdb(pdbs$id)
 
```


```{r}
# Calculate sequence identity
seqidentity(pdbs)

# Calculate RMSD
rmsd(pdbs)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)   

# Clustering  
hc <- hclust(as.dist(rd))
grps <- cutree(hc, k=3) 
 
# Plot results as dendrogram
hclustplot(hc, k=3)
```

# Section 5- PCA of adenylate kinase structures
```{r}
#getting pdb file for adenylate cyclase
aa <- get.seq("1ake_A")
aa

```

Search the PDB database
```{r, cache=TRUE}
b <- blast.pdb(aa)
```

```{r}
hits <- plot(b)
```

## PCA
```{r}
#find invariant core
core <- core.find(pdbs)

#superimpose all structures to core

```

