---
title: "class14"
author: "Vaibhav Menon"
date: "5/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import countData and colData into R
```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors=FALSE)
metadata <-read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)

```

Id'ing and calculating mean of control samples
```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/length(control$id) 
names(control.mean) <- counts$ensgene
head(control.mean)
```

Id'ing and calculating mean of treated samples
```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums( counts[ ,treated$id])/length(treated$id)
names(treated.mean) <- treated$ensgene
head(treated.mean)
```

Combining meancount data for bookkeeping purposes
```{r}
meancounts <- data.frame(control.mean, treated.mean)
colSums(meancounts)
```

Scatter plot mean treated vs mean control
```{r}
plot.default(meancounts$control.mean, meancounts$treated.mean, log="xy", xlab="log control", ylab="log treated")
```

Finding differentially expressed genes
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

Getting rid of NaN's and inf's
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)
to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```
Setting two-fold threshold for differnetial expression
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
paste("Up: ", sum(up.ind))
paste("Down: ", sum(down.ind))
```
```{r}
anno <- read.csv("data/annotables_grch38.csv")
results <- merge(mycounts, anno, by.x="row.names" ,by.y="ensgene" )
head(results)
```

# Annotation via packages
Loading packages
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

Annotation
```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")

mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
mycounts$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
head(mycounts)
```
# DEseq2 analysis
Load the package
```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

Getting results
```{r}
res <- results(dds)
res
```
```{r}
summary(res)
```

Reorder results table by the smallest p-value
```{r}
resOrdered <- res[order(res$pvalue),]
```

customize Alpha value
```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

Determine significant differential expression
```{r}
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)

resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

Annotating results dataframe for alpha=0.01
```{r}
res01 <- results(dds, alpha=0.01)
summary(res01)

resSig01 <- merge(resSig01, anno, by.x="row.names" ,by.y="ensgene" )
head(resSig01)
```

Ordered by smallest p value
```{r}
ord <- order( resSig01$padj )
#View(res01[ord,])
head(resSig01[ord,])
```

```{r}
write.csv(resSig01[ord,], "signif01_results.csv")
```

# Data Visulization
```{r}
i <- grep("CRISPLD2", resSig01$symbol)
resSig01[i,]
```
```{r}
rownames(resSig01[i,])
```

```{r}
plotCounts(dds, gene="ENSG00000103196", intgroup="dex")
```

Return the data
```{r}
d <- plotCounts(dds, gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
head(d)
```

Boxplot, default and ggplot
```{r}
boxplot(count ~ dex , data=d)
library(ggplot2)
ggplot(d, aes(dex, count)) + geom_boxplot(aes(fill=dex)) + scale_y_log10() + ggtitle("CRISPLD2")
```

MA & Volcano plots
```{r}
res$sig <- res$padj<0.05

# How many of each?
table(res$sig)

sum(is.na(res$sig))
```

MA Plot
```{r}
plotMA(res, ylim=c(-2,2))
```

Get rid of noise from 2fold log changes 
```{r}
resLFC <- lfcShrink(dds, coef=2)
resLFC
plotMA(resLFC, ylim=c(-2,2))
```

Volcano Plots
```{r}
ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```

