---
title: "class13"
author: "Vaibhav Menon"
date: "5/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1000 genome project data aanysis

downloaded genotype data from ensemble for Asthma associated SNP rs8067378
See: MXL Mexican ancestry in Los Angeles, California
```{r}
#read csv file
genotype <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head (genotype)
```
Find how many of each genotype there are in this dataset
```{r}
genotype.summary <- table(genotype[,2])
genotype.summary
```
Proportion of each genotype
```{r}
genotype.summary/nrow(genotype)*100
```

## RNA-seq
Look more closely at sample HG00109 with G|G genotype for this SNP (from GBR population)

### side note on fastq quality scores
```{r}
#install.packages("seqinr")
#install.packages("gtools")
library(seqinr)
library (gtools)
phred <- asc(s2c("DDDDCDDEEDEDFFFFHHHHHHJJIIJIGIJJJJIJJIJIHFD;JJIJJJIJIIJJJJJIJJHHHHHFFFFFCCC"))-33
phred
```
```{r}
prob <- 10**(-phred/10)
prob
```

## Population Scale Analysis
```{r}
genotype.table <- read.table("rs8067378_ENSG00000172057.6.txt")
summary(genotype.table)
```

Look at data
```{r}
table(genotype.table$geno)
```
How many samples?
```{r}
nrow(genotype.table)
```

```{r}
# boxplot
plot <- boxplot(genotype.table)
plot
```


